### 2011-12-08 doc/theses/mikula/analysis_of_recent_approaches.tex
Master thesis: In progress.

\chapter{Analysis of Recent Approaches} \label{chapter_analysis_of_recent_approaches}
Existing approaches of XML schema inference can be classified by several criteria. A basic division is by the language the resulting schema is written in. Commonly used languages are DTD and XML Schema.

According to \cite{Mlynkova:2008:AAX:1494650.1495496}, the type of the inference method can be divided into \emph{heuristic} and \emph{grammar-inferring}. Heuristic methods (\cite{Chidlovskii:2002:SEX:544220.544288, Garofalakis:2000:XSE:342009.335409, Moh:2000:RSW:336597.336638, Wong03onstructural, Vosta:2008:EAC:1802514.1802522}) are based on experience with manual construction of schemas, motivated by real-world usages of XML schema, and their result commonly does not belong to any class of grammar. Two of these approaches are described in Section \ref{section_xtract} and Section \ref{section_even_an_ant_can_create_an_xsd}.

On the contrary, the grammar-inferring methods (\cite{Ahonen1996GeneratingGrammars, Bex:2006:ICD:1182635.1164139, Bex:2007:IXS:1325851.1325964, Hegewald:2006:XES:1129755.1130124, Min:2003:EES:639473.639475, Fernau:2001:LXG:645744.667236}) are based on theoretical knowledge of automata and their results belong to a particular class of languages. Thus, these methods guarantee specific characteristics of the results.

Another important criterion is the type of input data. Most of the approaches process XML documents as the input of the inference process and the documents are supposed to be valid against the resulting schema. Besides approaches exploiting XML data, approaches that utilize other or additional sources may be developed. An approach utilizing XML data along with an obsolete XML schema is described in \cite{Mlynkova:2009:IXS:1862681.1862693}. However, the most significant approaches in terms of this work are those, utilizing operations over XML data. According to our best knowledge at the time of writing, there is just one approach of this category, described in \cite{Necasky:2009:DXK:1529282.1529414}. It utilizes a set of XQuery queries to discover keys and foreign keys.

\section{Common Caracteristics}
The process of inference commonly used by a significant number of approaches is summarized in \cite{Mlynkova:2008:AAX:1494650.1495496} as the following one: For each occurrence of element $e$ from the input XML documents and its subelements $e_1, e_2, ..., e_k$ a production $e \rightarrow e_1 e_2 ... e_k$ is constructed. The productions form so-called \emph{initial grammar} (IG). For each element type the productions are then merged, simplified and generalized using various methods and criteria. A common approach is so-called \emph{merging state algorithm}, where a \emph{prefix tree automaton} (PTA) is built from the productions of the same element type and the automaton is generalized via merging of its states. Finally, the generalized automaton/grammar is expressed in syntax of the respective XML schema language.
\todo[inline]{Tento odstavec je doslovne prebraty z uvedeneho clanku. Mozem to takto pouzit?}

@@

\subsection{Generalization}
Purpose of generalization is to create a set of DTD candidates - schemata that cover fractions of the input XML data. In the last step, this set will be used to compose a (sub)optimal result with respect to a trade-off between its preciseness and conciseness. Therefore, it is desirable to create DTD canditates with various degree of these two characteristics.

@@

For instance, paper \cite{Garofalakis:2000:XSE:342009.335409} introduces the following example: Sequences \texttt{abab} and \texttt{bbbe} are generalized to \texttt{(ab)*}, \texttt{(a|b)*}, \texttt{b*e}.
\todo[inline]{Mozem takto pouzivat prevzate priklady?}

\subsection{Factoring}
Factoring is a process of creating a new DTD canditate from two or more DTD canditates, decresing their summed size without modifications in their semantics. An aim of this step is to decrease an MDL cost of DTD candidates calculated in the MDL step and thus refine the process of construction of the resulting DTD.

@@

Alike the generalization step, also in this step the set of possible factored DTD is huge and the authors propose certain heuristics to make the factorization effective.

@@

\todo[inline]{Detailnejsie? Priklad?}

@@

\begin{itemize}
\item Distinguishing elements with the same name but different semantics.
\item Improvements of algorithms adopted from the previous works.
\item Incorporating inference of an unordered sequence.
\item Creating a result in XML Schema language.
\end{itemize}

\subsection{Clustering of Elements}
This phase clusters elements on a basis of their context and structure. It is done by creating tree structures for each input XML documents where vertices represent elements and attributes and an edge between two vertices expresses their parent-child relation. These trees and their subtrees are then compared using an imposed tree similarity measure to find elements with the same semantics.

\subsection{Schema Generalization}
For each cluster, a trivial schema is created, which is then generalized to achieve a reasonable result. In a search for the optimal schema, \emph{Ant Colony Optimization (ACO)} heuristic is incorporated. The idea behind the ACO heuristic is that a set of artifical ants is searching a space of possible solutions, each ant given a subspace of the space to find a local suboptimum. An ant 
is performing steps (schema modifications), dying after a predefined number of steps and providing an information - positive feedback - on quality of a solution found. The search is performed in a defined count of interations (or it stops if a good enough solution is found) and the positive feedback from one iteration is used to find better results in the following iterations. Every step of an ant represents a modification of a schema, in particular, a merge of states in a corresponding PTA.

@@

\subsection{Result in XSD}
Unlike the majority of recent approaches, this methods creates its result in XML Schema language (XSD). Comparing to DTD, XSD is a newer language and it provides greater expression possibilities. Authors of this method focused on inferring elements with the same name but different context and the unordered sequences which can be in XSD expressed by \texttt{xs:all} construct. Elements with the same name but different context cannot be expressed in DTD and, however, the unordered sequence can be also expressed in DTD as alternations of ordered sequences, such expression in not practical nor well human-readable.

\section{On Inference of XML Schema with the Knowledge of an Obsolete One}
An aim of approach described in \cite{Mlynkova:2009:IXS:1862681.1862693} is to exploit an obsolete XML schema as an additional input information to infer a new schema more efficiently. A XML schema can become obsolete due to changes in a set of XML documents, without capturing these changes in the schema. Thus, the schema becames outdated and according to the paper, this case is quite common.

@@

And the algorithm consists of two independent steps:

@@

In the fist step the input schema is corrected to conform to the whole set of input XML documents. This is done by creating a PTA \todo{ref na def} for each production extracted from the input schema and merging it with a respective production from the initial grammar \todo{ref na def}, involving ACO and MDL heuristics.

In the second step, assuming a correct schema it is wanted to specialize the regular expressions involved in the schema with regard to the XML documents, resulting in more precise and readable schema. Optional substeps are pruning of unused schema fragments, correction of lower and upper bounds of occurrences, transcription of operators to a more restrictive but simplier form, if this transcription preserves the validity, and refactorization to improve readability.

\section{Discovering XML Keys and Foreign Keys in Queries}
This method is described in paper \cite{Necasky:2009:DXK:1529282.1529414}. It tries to improve automatic XML schema inference by discovering keys and foreign keys from a set of XQuery queries. Just the queries are utilized in inference, no XML data are used. The output of this method is a set of keys and foreign keys that can be captured using XML Schema \textbf{key}, \textbf{keyref} and \textbf{unique} constructs.

@@

The method is based on assumption that each join is done via key/foreign key pair. It means it is supposed that $L_1$ is a key of the elements $S_1$ and $L_2$ is it's respective foreign key or vice versa.

The author describe two possible cases:

\begin{enumerate}
\renewcommand{\theenumi}{(O\arabic{enumi})}
\renewcommand{\labelenumi}{\theenumi}
\item $L_1$ is a key of the elements $S_1$, $L_2$ is a respective foreign key and it itself is not a key of the elements $S_2$.
\item $L_2$ is a key of the elements $S_2$, $L_1$ is a respective foreign key and it cannot be decided whether $L_1$ is a key of the elements $S_1$ or not.
\end{enumerate}

@@

The pattern occurrence is considered of case (O1) if it is the for join pattern (weight: 1), if aggregation function \texttt{avg}, \texttt{min}, \texttt{max} or \texttt{sum} is applied on a return path (weight: 1) or if aggregation function \texttt{count} is applied on a return path (weight: 0.75). \todo {definovat return path?}

@@

The assumption the method is based on may not be fulfilled for every join in a supplied set of queries. A key $K$ may be inferred from some query and processing of another query may result to denial of $K$ as a key. \todo[inline]{priklad?}

@@

\subsection{Conclusion}
The output of the method is a list of scored keys and for each key a list foreign keys referencing the key. Since the method deals only with the inference of keys, it is not an overall method of the XML schema inference. It is meant to be used in a collaboration with other schema-inferring methods to refine their results.

@@

\section{Summary}
\begin{table}
  \begin{threeparttable}
  \footnotesize \begin{tabular}{|c|c|c|c|}
    \hline
    Method name and/or paper & Input & Output & Year \\ \hline \hline
    \cite{Ahonen1996GeneratingGrammars} & XML documents & DTD & 1996 \\ \hline
    XTRACT \cite{Garofalakis:2000:XSE:342009.335409} & XML documents & DTD & 2000 \\ \hline
    DTD-miner \cite{Moh:2000:RSW:336597.336638} & XML documents & DTD & 2000 \\ \hline
    \cite{Fernau:2001:LXG:645744.667236} & XML documents & DTD & 2001 \\ \hline
    ECFG \cite{Chidlovskii:2002:SEX:544220.544288} & XML documents & XSD & 2002 \\ \hline
    sk-ANT \cite{Wong03onstructural} & XML documents & DTD & 2003 \\ \hline
    \cite {Min:2003:EES:639473.639475} & XML documents & DTD & 2003 \\ \hline
    \cite{Bex:2006:ICD:1182635.1164139} & XML documents & DTD & 2006 \\ \hline
    XStruct \cite {Hegewald:2006:XES:1129755.1130124} & XML documents & XSD & 2006 \\ \hline
    \cite{Bex:2007:IXS:1325851.1325964} & XML documents & XSD & 2007 \\ \hline
    SchemaMiner \cite{Vosta:2008:EAC:1802514.1802522} & XML documents & XSD & 2008 \\ \hline
    \cite{Mlynkova:2009:IXS:1862681.1862693} & XML documents, XML schema & DTD, XSD & 2009 \\ \hline
    \cite{Necasky:2009:DXK:1529282.1529414} & XQuery queries & --\tnote{1} & 2009 \\ \hline    
  \end{tabular}
  \todo[inline]{doplnit}
  \begin{tablenotes}
	\item [1] Result of this method is a list of discovered keys and foreign keys.
  \end{tablenotes}
  \caption{Summary of Recent Approaches}
  \label{table_summary_of_recent_approaches}
  \end{threeparttable}
\end{table}

As shown in Table \ref{table_summary_of_recent_approaches}, most of the recent approaches of XML schema inference are base on utilization of XML documents. These incorporates various verified methods and the newer approaches often improve the older ones and/or combine them together to achieve better results.

Lately, several approaches that utilize other input sources have been proposed. Paper \cite{Mlynkova:2009:IXS:1862681.1862693} introduces a method that utilizes an XML schema besides XML documents. The method described in \cite{Necasky:2009:DXK:1529282.1529414} utilizes XQuery queries, however, the result of this method is a list of discovered keys and foreign keys, not an XML schema. Also, to our best knowledge, implementation that combines this method with another approach to get the XML schema has not been proposed yet. \todo{Toto je pravda, naozaj ta implementacia neexistuje?}

>>>>

\chapter{Analysis of Recent Approaches} \label{chapter_analysis_of_recent_approaches}
Existing approaches to XML schema inference can be classified using several criteria. A basic classification is based on the language the resulting schema is written in. Commonly used languages are DTD and XML Schema.

According to \cite{Mlynkova:2008:AAX:1494650.1495496}, the type of the inference method can be divided into \emph{heuristic} and \emph{grammar-inferring}. Heuristic methods \cite{Chidlovskii:2002:SEX:544220.544288, Garofalakis:2000:XSE:342009.335409, Moh:2000:RSW:336597.336638, Wong03onstructural, Vosta:2008:EAC:1802514.1802522} are based on experience with manual construction of schemas, motivated by real-world usages of XML schema, and their result commonly does not belong to any class of grammar. Two of these approaches are described in Section \ref{section_xtract} and Section \ref{section_even_an_ant_can_create_an_xsd}. \todo{chce to větu proč zrovna dva}

On the contrary, the grammar-inferring methods \cite{Ahonen1996GeneratingGrammars, Bex:2006:ICD:1182635.1164139, Bex:2007:IXS:1325851.1325964, Hegewald:2006:XES:1129755.1130124, Min:2003:EES:639473.639475, Fernau:2001:LXG:645744.667236} are based on theoretical knowledge of automata and their results belong to a particular class of languages. Thus, these methods guarantee specific characteristics of the results.

Another important criterion is the type of input data. Most of the approaches process XML documents as the input of the inference process and the documents are supposed to be valid against the resulting schema. Besides approaches exploiting XML data, approaches that utilize other or additional sources may be developed. An approach utilizing XML data along with an obsolete XML schema is described in \cite{Mlynkova:2009:IXS:1862681.1862693}. However, the most significant approaches in terms of this work are those utilizing operations over XML data. According to our best knowledge at the time of writing, there is just one approach of this category, described in \cite{Necasky:2009:DXK:1529282.1529414}. It utilizes a set of XQuery queries to discover keys and foreign keys.

\section{Common Caracteristics}
The process of XML schema inference commonly used by a significant number of approaches is summarized in \cite{Mlynkova:2008:AAX:1494650.1495496} as the following one: For each occurrence of element $e$ from the input XML documents and its subelements $e_1, e_2, ..., e_k$ a production $e \rightarrow e_1 e_2 ... e_k$ is constructed. The productions form so-called \emph{initial grammar} (IG). For each element type the productions are then merged, simplified and generalized using various methods and criteria. A common approach is so-called \emph{merging state algorithm}, where a \emph{prefix tree automaton} (PTA) is built from the productions of the same element type and the automaton is generalized via merging of its states. Finally, the generalized automaton/grammar is expressed in syntax of the respective XML schema language.
\todo[inline]{Q: Tento odstavec je doslovne prebraty z uvedeneho clanku. Mozem to takto pouzit? A: ano, jen by pak neměla být výrazně jiná úroveň angličtiny}

@@

\subsection{Generalization}
The purpose of generalization is to create a set of DTD candidates - schemata that cover fractions of the input XML data. In the last step, this set will be used to compose a (sub)optimal result with respect to a trade-off between its preciseness and conciseness. Therefore, it is desirable to create DTD canditates with various degrees of these two characteristics.

@@

For instance, paper \cite{Garofalakis:2000:XSE:342009.335409} introduces the following example: Sequences \texttt{abab} and \texttt{bbbe} are generalized to \texttt{(ab)*}, \texttt{(a|b)*}, \texttt{b*e}.

\subsection{Factoring}
Factoring is a process of creating a new DTD canditate from two or more DTD canditates, decresing their summed size without modifications in their semantics. The aim of this step is to decrease the MDL cost of DTD candidates calculated in the MDL step and thus refine the process of construction of the resulting DTD.

@@

Alike the generalization step, also in this step the set of possible factored DTDs is huge and the authors propose certain heuristics to make the factorization effective.

@@

@@

\begin{itemize}
\item Distinguishing elements with the same name but different context.
\item Improvements of algorithms adopted from the previous works.
\item Incorporating inference of an unordered sequence.
\item Creating a result in the XML Schema language.
\end{itemize}

\subsection{Clustering of Elements}
This phase clusters elements on the basis of their context and structure. It is done by creating tree structures for each input XML document where vertices represent elements and attributes and an edge between two vertices expresses their parent-child relationship. These trees and their subtrees are then compared using an imposed tree similarity measure to find elements with the same semantics.

\subsection{Schema Generalization}
For each cluster, a trivial schema is created, which is then generalized to achieve a reasonable result. In a search for the optimal schema, \emph{Ant Colony Optimization (ACO)} heuristic is incorporated. The idea behind the ACO heuristic is that a set of artifical ants is searching a space of possible solutions, each ant given a subspace of the space to find a local suboptimum. An ant is performing steps (schema modifications), dying after a predefined number of steps and providing an information - positive feedback - on the quality of a solution found. The search is performed in a defined number of interations (or it stops if a good enough solution is found) and the positive feedback from one iteration is used to find better results in the following iterations. Every step of an ant represents a modification of a schema, in particular, a merge of states in a corresponding PTA.

@@

\subsection{Result in XSD}
Unlike the majority of recent approaches, this methods creates its result in XML Schema. Contrary to DTD, XSD is a newer language and it provides greater expressive possibilities. The authors of this method focused on inferring elements with the same name but different context and the unordered sequences which can be in XSD expressed by \texttt{xs:all} construct. Elements with the same name but different context cannot be expressed in DTD and, although, the unordered sequence can be also expressed in DTD as alternations of ordered sequences, such expression in not practical nor well human-readable.

\section{On Inference of XML Schema with the Knowledge of an Obsolete One}
The aim of approach described in \cite{Mlynkova:2009:IXS:1862681.1862693} is to exploit an obsolete XML schema as an additional input information to infer a new schema more efficiently. A XML schema can become obsolete due to changes in a set of XML documents, without capturing these changes in the schema. Thus, the schema becomes outdated and according to the paper, this case is quite common.

@@

The algorithm consists of two independent steps:

@@

In the fist step the input schema is corrected to conform to the whole set of input XML documents. This is done by creating a PTA for each production extracted from the input schema and merging it with a respective production from the initial grammar, involving ACO and MDL heuristics.

\todo{Tato věta hodně divná.} In the second step, assuming a correct schema it is wanted to specialize the regular expressions involved in the schema with regard to the XML documents, resulting in more precise and readable schema. Optional substeps are pruning of unused schema fragments, correction of lower and upper bounds of occurrences, transcription of operators to a more restrictive but simplier form, if this transcription preserves the validity, and refactorization to improve readability.

\section{Discovering XML Keys and Foreign Keys in Queries}
The method described in paper \cite{Necasky:2009:DXK:1529282.1529414} improves automatic XML schema inference by discovering keys and foreign keys from a set of XQuery queries. Just the queries are utilized in inference, no XML data are used. The output of this method is a set of keys and foreign keys that can be captured using XML Schema \textbf{key}, \textbf{keyref} and \textbf{unique} constructs.

@@

The method is based on assumption that each join is done via key/foreign key pair. It means it is supposed that $L_1$ is a key of the elements $S_1$ and $L_2$ is its respective foreign key or vice versa.

The authors describe two possible cases:

\begin{enumerate}
\renewcommand{\theenumi}{(O\arabic{enumi})}
\renewcommand{\labelenumi}{\theenumi}
\item $L_1$ is a key of the elements in $S_1$, $L_2$ is a respective foreign key and it itself is not a key of the elements in $S_2$.
\item $L_2$ is a key of the elements in $S_2$, $L_1$ is a respective foreign key and it cannot be decided whether $L_1$ is a key of the elements in $S_1$ or not.
\end{enumerate}

@@

The pattern occurrence is considered of case (O1) if it is the for join pattern (weight: 1), if aggregation function \texttt{avg}, \texttt{min}, \texttt{max} or \texttt{sum} is applied on a return path (weight: 1) or if aggregation function \texttt{count} is applied on a return path (weight: 0.75). \todo {Q: definovat return path? A: no, možná spíš jednu větu co to zhruba je}

@@

The assumption the method is based on may not be fulfilled for every join in a supplied set of queries. A key $K$ may be inferred from some query and processing of another query may result to denial of $K$ as a key.

@@

\subsection{Conclusion}
The output of the method is a list of scored keys and for each key a list foreign keys referencing the key. Since the method deals only with the inference of keys, it is not an overall method of the XML schema inference. It is meant to be used in collaboration with other schema-inferring methods to refine their results.

@@

\section{Summary}
\begin{table}
  \begin{threeparttable}
  \footnotesize \begin{tabular}{|c|c|c|c|}
    \hline
    Method name and/or paper & Input & Output & Year \\ \hline \hline
    \cite{Ahonen1996GeneratingGrammars} & XML documents & DTD & 1996 \\ \hline
    XTRACT \cite{Garofalakis:2000:XSE:342009.335409} & XML documents & DTD & 2000 \\ \hline
    DTD-miner \cite{Moh:2000:RSW:336597.336638} & XML documents & DTD & 2000 \\ \hline
    \cite{Fernau:2001:LXG:645744.667236} & XML documents & DTD & 2001 \\ \hline
    ECFG \cite{Chidlovskii:2002:SEX:544220.544288} & XML documents & XSD & 2002 \\ \hline
    sk-ANT \cite{Wong03onstructural} & XML documents & DTD & 2003 \\ \hline
    \cite {Min:2003:EES:639473.639475} & XML documents & DTD & 2003 \\ \hline
    \cite{Bex:2006:ICD:1182635.1164139} & XML documents & DTD & 2006 \\ \hline
    XStruct \cite {Hegewald:2006:XES:1129755.1130124} & XML documents & XSD & 2006 \\ \hline
    \cite{Bex:2007:IXS:1325851.1325964} & XML documents & XSD & 2007 \\ \hline
    SchemaMiner \cite{Vosta:2008:EAC:1802514.1802522} & XML documents & XSD & 2008 \\ \hline
    \cite{Mlynkova:2009:IXS:1862681.1862693} & XML documents, XML schema & DTD, XSD & 2009 \\ \hline
    \cite{Necasky:2009:DXK:1529282.1529414} & XQuery queries & --\tnote{1} & 2009 \\ \hline    
  \end{tabular}
  \todo[inline]{doplnit}
  \begin{tablenotes}
	\item [1] The result of this method is a list of discovered keys and foreign keys.
  \end{tablenotes}
  \caption{Summary of Recent Approaches}
  \label{table_summary_of_recent_approaches}
  \end{threeparttable}
\end{table}

As shown in Table \ref{table_summary_of_recent_approaches}, most of the recent approaches of XML schema inference are based on utilization of XML documents. These incorporate various verified methods and the newer approaches often improve the older ones and/or combine them together to achieve better results.

Lately, several approaches that utilize other input sources have been proposed. Paper \cite{Mlynkova:2009:IXS:1862681.1862693} introduces a method that utilizes an XML schema besides XML documents. The method described in \cite{Necasky:2009:DXK:1529282.1529414} utilizes XQuery queries, however, the result of this method is a list of discovered keys and foreign keys, not an XML schema. Also, to our best knowledge, implementation that combines this method with another approach to get the XML schema has not been proposed yet.

### 2011-12-08 doc/theses/mikula/introduction.tex
Master thesis: In progress.

\chapter{Introduction}
Extensible Markup Language (XML) \cite{Bray:08:EML} has become a popular standard for data representation and exchange. Its prevalence is based on its simplicity and at the same time its flexibility and express power.

With XML used to represent data themselves, it is often needed or convenient to somehow specify a structure of the represented data, their format, internal relations or restrictions, etc. In order to this need, several so-called XML schema languages (XML schemata) have been created. The most used of them are Document Type Definition (DTD) \cite{Bray:08:EML} and XML Schema (or XSD as XML Schema Definition) \cite{Walmsley:04:XSP, Thompson:04:XSP, Malhotra:04:XSP} proposed by The World Wide Web Consorcium (W3C).

Nevertheless an encouragement to use an XML schema along with an XML data representation, in practise it is done sparsely. Commonly, XML documents are not assigned with their respective XML schema at all or the schema is outdated due to modifications done to a structure of data without updating the schema.

Recently in reaction to this situation, significant number of approaches dealing with an automatic construction of an XML schema have been proposed. An aim is to exploit a provided set of XML documents and infer an XML schema, so that the XML documents are valid against it. In addition, the inferred schema should be reasonable in views of human-readability, preciseness and conciseness. Most of published approaches of an XML schema inference are of this type - input is a set of XML documents. They are based on various ideas and can be classified by several aspects as discussed in Chapter \ref{chapter_analysis_of_recent_approaches}.

Besides the mentioned type, some approaches that utilize other or additional sources has been developed. If there are available sources like an out-dated XML schema, operations upon the XML data such as a set of XQuery \todo{ref} queries, or other, these sources can be exploited to refine the process of inference. The refinement can be achieved in various aspects such as decreasing the speed of the process, getting a more precise, more concise or more readable result or inference of some statements about the data which cannot be (easily) extracted from the data themselves.

Recently, the main effort has been focused on a research of the approaches that utilize XML documents, and thus, there are only few approaches of the latter type (also discussed in Chapter \ref{chapter_analysis_of_recent_approaches}), leaving a wide space for a possible future research and improvements.

\section{Aim of this Work}
\todo[inline]{Toto je skopirovane zo sis-u zo zadania prace, moze to tu byt?}
The aim of this work is a research on the problem of inference of an XML schema for the given set of XML data in a situation when we are provided also with a set of related operations (XML queries, XSLT scripts etc.).

Firstly, it is necessary to analyze existing inference solutions in general and to discuss their advantages and disadvantages. The core of the work is identification and discussion of information that can be extracted from a given set of XML operations and how they can be exploited to achieve more precise and realistic XML schema.

The result of the work will be a proposal of own approach involving the improvements, its implementation and suitable experiments that will show its advantages.

\section{Structure of this Work}
\todo[inline]{TODO}

>>>>

\chapter{Introduction}
The Extensible Markup Language (XML) \cite{Bray:08:EML} has become a popular standard for data representation and exchange. Its prevalence is based on its simplicity and, at the same time, its flexibility and expressive power.

With XML used to represent data themselves, it is often needed or convenient to somehow specify a structure of the represented data, their format, internal relations or restrictions, etc. In order to this need, several so-called XML schema languages (XML schemata) have been created. The most used of them are Document Type Definition (DTD) \cite{Bray:08:EML} and XML Schema (or XSD as XML Schema Definition) \cite{Walmsley:04:XSP, Thompson:04:XSP, Malhotra:04:XSP} proposed by the World Wide Web Consorcium (W3C).

Despite an conveyed encouragement to use an XML schema along with an XML data representation, in practise it is done sparsely. Commonly, XML documents are not assigned with their respective XML schema at all or the schema is outdated due to modifications done to a structure of data without updating the schema. \todo{ref na clanok s tvrdenim}

Recently in reaction to this situation, a significant number of approaches dealing with an automatic construction of an XML schema has been proposed. An aim is to exploit a provided set of XML documents and infer an XML schema, so that the XML documents are valid against it. In addition, the inferred schema should be reasonable in views of human-readability, preciseness and conciseness. Most of published approaches to XML schema inference are of this type - the input is a set of XML documents. They are based on various ideas and can be classified by several aspects as discussed in Chapter \ref{chapter_analysis_of_recent_approaches}.

Besides the mentioned type, some approaches that utilize other or additional sources have been developed \todo{reference, např. DP Klempa}. If there are available sources like an out-dated XML schema, operations upon the XML data such as a set of XQuery \todo{ref} queries, or other, these sources can be exploited to refine the process of inference. The refinement can be achieved in various aspects such as increasing the speed of the process, getting a more precise, more concise or more readable result or inference of statements about the data which cannot be (easily) extracted from the data themselves.

Recently, the main effort has been focused on a research of the approaches that utilize XML documents, and thus, there are only few approaches of the latter type (as also discussed in Chapter \ref{chapter_analysis_of_recent_approaches}), leaving a wide space for a possible future research and improvements.

\section{Aim of this Work}
Our aim is to perform a research on the problem of inference of an XML schema for the given set of XML data in a situation when we are provided also with a set of related operations (XML queries, XSLT scripts \todo{ref} etc.).

Firstly, we indend to analyze existing inference solutions in general and to discuss their advantages and disadvantages. Then, we try to identificate and discuss information that can be extracted from a given set of XML operations and how they can be exploited to achieve more precise and realistic XML schema.

The work should result into a proposal of own approach involving the improvements, its implementation and suitable experiments that will show its advantages.

\section{Structure of the Work}
\todo[inline]{TODO}

### 2011-12-08 doc/theses/mikula/proposed_algorithm.tex
Master thesis: In progress.

How are the types inferred from the expression types? We do not exploit all expressions. Only expressions of a particular type are exploited. Specifically, an expression has to contain some $pathType$ subexpression $E$ (expression representing a certain element or attribute or set of elements or attributes). In the following text, the set represented by expression $E$ is denoted $S$. Another requirement is the expression has to be either a function call or an arithmetic operation \todo{Nejake ine ako napr type constructor, cast as? Princip je stale rovnaky}.

\todo[inline]{Zlucovanie odvodenych tvrdeni}
\todo[inline]{Co ked mame path expression s predikatmi?}

>>>>

How are the types inferred from the expression types? We do not exploit all expressions. Only expressions of a particular type are exploited. Specifically, an expression has to contain an subexpression $E$ of $pathType$ type (expression representing a certain element or attribute or set of elements or attributes). In the following text, the set represented by expression $E$ is denoted $S$. Another requirement is the expression has to be either a function call or an arithmetic operation \todo{Nejake ine ako napr type constructor, cast as? Princip je stale rovnaky}.

The output of this step is a set of couples ($pathType$, XML Schema built-in type).

\todo[inline]{Zlucovanie odvodenych tvrdeni.}
\todo[inline]{Path expressions s predikatmi.}

### 2011-12-08 doc/theses/mikula/used_technologies_and_definitions.tex
Master thesis: In progress.

\section{XML schema}
An XML schema refers to a description of an XML document in terms of its structure and various constraints. Commonly, the XML schema describes element and attribute names, their parent-child relations, their order and a type of their content. Other constraints often expressed in the XML schema are restrictions on numbers of occurrences of elements, specification of (non-)obligatory attributes, uniqueness and specification of keys.

@@

A validity of an XML document againt its XML schema expresses wheather the document is well-formed \todo{link} and at the same time wheather it conforms to the XML schema.

@@

The content of an element can be denoted by \emph{EMPTY} for an empty element, \emph{ANY} for any content, \emph{(\#PCDATA)} allowing only textual content (without any other subelements), or specified using regular expressions. Names of subelement are combined using the well-known operators (\emph{|}, \emph{+}, \emph{*}, \emph{?} and \emph{,}(comma)). To express the mixed content \emph{\#PCDATA} can be used in an alternation list with the subelement names and this alternation has to be enclosed in \emph{*} operator.

>>>>

\section{XML Schema}
An XML schema refers to a description of an XML document in terms of its structure and various constraints. Commonly, the XML schema describes element and attribute names, their parent-child relations, their order and a type of their content. Other constraints often expressed in the XML schema are restrictions on numbers of occurrences of elements, specification of (non-)obligatory attributes, uniqueness and specification of keys.

@@

A validity of an XML document againt its XML schema expresses wheather the document is well-formed \todo{link}, and at the same time, wheather it conforms to the XML schema.

@@

The content of an element can be denoted by \emph{EMPTY} for an empty element, \emph{ANY} for any content, \emph{(\#PCDATA)} allowing only textual content (without any other subelements), or specified using regular expressions. Names of subelements are combined using operators (\emph{|}, \emph{+}, \emph{*}, \emph{?} and \emph{,}(comma)). To express the mixed content \emph{\#PCDATA} can be used in an alternation list with the subelement names and this alternation has to be enclosed in \emph{*} operator.
